
\documentclass[]{bytedance_seed}

% single-column: \documentclass[]{bytedance_seed}, 
%Please prioritize using single-column。

% twocolumn: \documentclass[twocolumn]{bytedance_seed}

\usepackage[toc,page,header]{appendix}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{minitoc}
\usepackage{amssymb}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage[utf8]{inputenc}  % for UTF-8 support
\usepackage{float}  % for listing environment
\usepackage{minted}  % for code highlighting
\usepackage{textcomp}  % for special symbols
\usepackage{footmisc}
% table 

\definecolor{papergreen}{HTML}{008000} % 较暗的绿色
\definecolor{paperred}{HTML}{D00000}   % 较暗的红色

% 命令 \goodres{主要值}{变化值}
% 用于显示一个好的结果 (绿色, +)
\newcommand{\goodres}[2]{%
    #1 {\color{papergreen}\small (+#2)}%
}
% 命令 \badres{主要值}{变化值}
% 用于显示一个坏的结果 (红色, -)
\newcommand{\badres}[2]{%
    #1 {\color{paperred}\small (-#2)}%
}
% 用于显示一个中性或无对比的结果
\newcommand{\neutralres}[2]{%
    #1 {\color{gray}\small (#2)}%
}
%%%%%%%%%%%%%%%%%%%%

\newcommand{\bench}{\texttt{Our-Bench}\xspace}
\definecolor{navyblue}{HTML}{0071BC}
\definecolor{oai-green-200}{HTML}{D7EADF} % 浅绿色
\definecolor{oai-green-400}{HTML}{A6D3B8} % 中等绿色
\definecolor{oai-green-600}{HTML}{63AC86} % 深绿色
\definecolor{oai-gray-300}{HTML}{D9D9D9}  % 浅灰色
\definecolor{oai-gray-600}{HTML}{808080}  % 深灰色
\definecolor{L1_Purple}{HTML}{E6E0F8} % A light purple for L1
\definecolor{L2_Orange}{HTML}{FCEFD8} % A light orange/yellow for L2
\definecolor{L3_Green}{HTML}{EFF8E0}  % A light green for L3
\definecolor{L4_Blue}{HTML}{E0F0F8}   % A light blue for L4
%%%%%%%%%%%%%%%%%%%%

% \title{SpatialTree: How Spatial Abilities Branch Out in MLLMs}
\title{SpatialTree: How Spatial Abilities Branch Out in MLLMs}

\author[\blacktriangle,\bigstar,*]{Yuxi Xiao}
\author[\diamondsuit,\bigstar,*]{Longfei Li}
\author[\bigstar]{Shen Yan}
\author[\bigstar]{Xinhang Liu}
\author[\blacktriangle]{Sida Peng}
\authorbreak
\author[\diamondsuit]{Yunchao Wei}
\author[\blacktriangle]{Xiaowei Zhou}
\author[\bigstar, \dagger]{Bingyi Kang}

%论文单位请使用ByteDance Seed
\affiliation[\blacktriangle]{Zhejiang University}
\affiliation[\bigstar]{ByteDance Seed}
\affiliation[\diamondsuit]{Beijing Jiaotong University}


\contribution[\dagger]{Project Lead}
\contribution[*]{Equal Contribution}

\abstract{
Spatial abilities in multimodal LLMs manifest hierarchically, spanning multiple levels from perception, through reasoning, to agentic tasks in 3D environments. However, cross-level dependencies among these abilities and their potential to guide data and training strategies for systematically scaling up remain largely unexplored. To address this, we propose SpatialTree, a hierarchical taxonomy organizing spatial abilities from low-level perception (L1) through mental mapping (L2) and mental simulation (L3) to agentic competence (L4). Building on this taxonomy, we construct the first hierarchical, capability-centric benchmark using our Spatial Engine. Guided by correlation analyses on our benchmark, we perform supervised fine-tuning (SFT) and reinforcement learning (RL) at different ability levels to probe their inter-dependencies. We find that low-level abilities (L1) are largely independent but can transfer to higher-level abilities and exhibit synergy during SFT. Furthermore, reinforcement learning proves more effective when applied to higher-level tasks rather than low-level abilities (L1), resulting in improved overall spatial performance. We hope our taxonomy and findings will inspire further research to systematically branch out spatial abilities in MLLMs through dependency-aware training.
}

\date{\today}
\correspondence{Bingyi Kang}
% You can add additional info fields as follows 
\checkdata[Project Page]{\url{xxx}}

\begin{document}
\maketitle

%不需要目录就注释掉 注意目录不要和第一页放在一块 要有\newpage
%\newpage
%\tableofcontents
%\newpage

\input{sections/introduction}
\input{sections/relatedwork}
\input{sections/approach}
\input{sections/experiments}


\clearpage

\bibliographystyle{plainnat}
\bibliography{main}

\clearpage

\beginappendix

\input{sections/appendix}

\end{document}